---
title: "Reverse Engineering Project"
author: "Liam Bowman, Peter Riccio, Andrew Mollenauer"
date: "Date here"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this notebook, we are reverse engineering the story, [STORY NAME](STORY%20URL)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library (tidyverse)
library(lubridate)
library(janitor)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data.

```{r}
# Load required data

homicide_data <- read_csv('data/homicide_data.csv') |> rename (date = reported_date) |> mutate (date=ymd(date))

#baltimore_homicides |> group_by (year) |> summarise (total_homicide = sum(n())) |> arrange(desc(total_homicide))

```

## Sentences to Engineer

In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

-   **Sentence text**: “City police made an arrest in 41 percent of homicides in 2014; last year, the rate was just 27 percent, a 14 percentage point drop.”
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

baltimore_homicides <- homicide_data |> filter (str_detect("Baltimore", city)) |>
  mutate (year= lubridate:: year(date))

baltimore_homicides <- baltimore_homicides |> mutate(
  no_arrest = case_when (
    str_detect (disposition, "Open/No arrest") ~ "YES",
    str_detect (disposition, "Closed without arrest") ~ "YES",
    .default = NA 
  )
)

baltimore_2014_homicides <- baltimore_homicides |> filter (year == 2014)

baltimore_2014_homicides_arrest_rate <- baltimore_2014_homicides |> select(no_arrest) |> filter (no_arrest = TRUE) |> nrow()/ baltimore_2014_homicides *100

  


# Display results of code below this codeblock

```

\|

### Sentence 2

-   **Sentence text**: “Of 50 of the nation’s largest cities, Baltimore is one of 34 where police now make homicide arrests less often than in 2014, according to a Washington Post analysis.”
-   **Analysis summary**: We were able to get an output close to the Post's finding. The code we wrote showed 33 cities that make arrests less often than 2014. Also confirmed that Baltimore was one of them. Part of the issue, it seems, is that the base dataset is incomplete. New York City homicides, for example, only start showing up in 2016.

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

#homicide_data |> group_by (city, year(date)) |> filter(str_detect (disposition, "Closed by arrest"))|> summarise (total_homicides = sum(n()))|> arrange (desc(total_homicides))

homicide_arrests_2014 <- homicide_data |> filter (str_detect(year(date), "2014"))

homicide_arrests_2014 <- homicide_arrests_2014 |> filter(str_detect(disposition,"Closed by arrest")) |> group_by (city) |> summarise(arrests2014 = sum(n())) |> arrange(desc(arrests2014))

homicides_2014 <- homicide_data |> filter(str_detect(year(date), "2014")) |> group_by (city)|> summarise (homicides2014 = sum(n()))

homicides_and_arrests_2014 <- homicides_2014 |> full_join (homicide_arrests_2014, join_by(city))

homicides_and_arrests_2014 <- homicides_and_arrests_2014 |> mutate(
  arrest_rate_2014 = arrest_rate_2014 <- (arrests2014/homicides2014) * 100
)

homicide_arrests_2017 <- homicide_data |> filter(str_detect(year(date), "2017"))

homicide_arrests_2017 <- homicide_arrests_2017 |> filter(str_detect(disposition,"Closed by arrest")) |> group_by (city) |> summarise(arrests2017 = sum(n())) |> arrange(desc(arrests2017))

homicides_2017 <- homicide_data |> filter(str_detect(year(date), "2017")) |> group_by (city) |> summarise (homicides2017 = sum(n()))

homicides_and_arrests_2017 <- homicides_2017 |> full_join (homicide_arrests_2017, join_by(city))

homicides_and_arrests_2017 <- homicides_and_arrests_2017 |> mutate(
  arrest_rate_2017 = arrest_rate_2017 <- (arrests2017/homicides2017) *100)

combined_homicide_arrest_rates <- homicides_and_arrests_2014 |> inner_join(homicides_and_arrests_2017, join_by(city))

combined_homicide_arrest_rates <- combined_homicide_arrest_rates |> mutate(
  arrests_less_often = case_when(
    arrest_rate_2014 >= arrest_rate_2017 ~ "YES",
    arrest_rate_2014 < arrest_rate_2017 ~ "NO"
  )
)

combined_homicide_arrest_rates |> filter(str_detect(arrests_less_often, "YES")) |> summarise (cities_with_fewer_arrests = sum(n()))



```

### Sentence 3

-   **Sentence text**: “And while most cities saw their arrest rates drop gradually, Baltimore’s decline was sudden — plummeting 15 percentage points in 2015, after Gray’s death, the largest single-year drop for any city already solving less than half its homicides.”
-   **Analysis summary**: We were able to replicate the Post's analysis here. Baltimore police made arrests in 25.4% of homicides in 2015, compared with 40.8% in 2014. It was, indeed, a roughly 15% drop.

```{r}
# Put code to reverse engineer sentence here

baltimore_homicides <- homicide_data |> group_by (city)|> filter(str_detect(city, "Baltimore")) 

b2014_arrest <- baltimore_homicides |> 
  filter(str_detect(year(date), "2014")) |> 
  filter(str_detect(disposition, "Closed by arrest"))|> 
           summarise(arrests2014 = sum(n()))

b2014_homicides <- baltimore_homicides |> 
  filter(str_detect(year(date), "2014")) |> 
  summarise(homicides2014 = sum(n()))

b2014_arrests_and_homicides <- b2014_homicides |> full_join (b2014_arrest, join_by(city))

b2014_arrests_and_homicides <- b2014_arrests_and_homicides |> mutate(
  arrest_rate_2014 = arrest_rate_2014 <- (arrests2014/homicides2014) *100)


b2015_arrest <- baltimore_homicides |> 
  filter(str_detect(year(date), "2015"))|>
           filter(str_detect(disposition, "Closed by arrest")) |>
           summarise(arrests2015 = sum(n()))
         
         
b2015_homicides <- baltimore_homicides |> 
  filter(str_detect(year(date), "2015")) |>
  summarise (homicides2015 = sum(n()))

b2015_arrests_and_homicides <- b2015_homicides |> full_join(b2015_arrest, join_by(city))

b2015_arrests_and_homicides <- b2015_arrests_and_homicides |> mutate(
  arrest_rate_2015 = arrest_rate_2015 <- (arrests2015/homicides2015) *100)

baltimore_combined_years <- b2014_arrests_and_homicides |> inner_join (b2015_arrests_and_homicides, join_by (city))


baltimore_combined_years
         


# Display results of code below this codeblock

```

### Sentence 4

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-
