---
title: "lab_12"
author: "Mitchell Hang"
date: "2023-05-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

-   tidytext and our usual libraries

## Load libraries and establish settings

**Task** Create a codeblock and load appropriate packages and settings for this lab.

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse.
library(tidytext)
library(tidyverse)
library(dplyr)
library(janitor)
library(lubridate)
library(rvest)
```

## Questions

##### **Q1.** You've been assigned to report a story about the leading reasons that Maryland attorneys get sanctioned by the state for misconduct. The state [publishes lists of sanctions](https://www.courts.state.md.us/attygrievance/sanctions) that contain a short text description about the situation. Load the CSV file in the data folder containing records from fiscal year 2011 onwards. Make a list of unique words from the text column, then following the example in the pre_lab, remove common "stop words" from that list and create a list of the top 10 words containing the percentage of occurrences each word represents. What's the leading word in that answer and, broadly, what do you think the top 10 words describe?

```{r}
#Load the data
md_attorney_sanctions <- read_csv("data/md_attorney_sanctions.csv")
```

```{r}
#Create list of unique words
unique_words <- md_attorney_sanctions |> select(text) |>
  unnest_tokens(word, text)
View(unique_words)

```

```{r}
#Load stop words
unique_words |>
  count(word, sort = TRUE) |>
  top_n(25) |>
  mutate(word = reorder(word, n)) |>
  ggplot(aes(x = word, y = n)) +
  geom_col() +
  xlab(NULL) +
  coord_flip() +
      labs(x = "Count",
      y = "Unique words",
      title = "Count of unique words found in sanctions")

```

```{r}
#Remove stop words from unique words list
data("stop_words")

stop_words <- stop_words |> 
  add_row(word = "to") |> 
  add_row(word = "the") |> 
  add_row(word = "for") |>
  add_row(word = "and") |>
  add_row(word = "of") |>
  add_row(word = 'his') |> 
  add_row(word = 'a') |> 
  add_row(word = 'on') |> 
  add_row(word = 'with') |> 
  add_row(word = "a") |> 
  add_row(word = "that") |> 
  add_row(word = "by") |> 
  add_row(word = "attorney") |> 
  add_row(word = "clients")|>
  add_row(word = "her") |>
  add_row (word = "respondent") |>
  add_row (word = "client")
  

```

```{r}
unique_words |>
  anti_join(stop_words) |>
  group_by(word) |>
  tally(sort=TRUE) |>
  mutate(percent = (n/sum(n))*100) |>
  top_n(10)
```

**A1. By a large margin, "failing" is the most commonly occurring word in attorney sanctions. Which makes sense: failing to do your job is the whole reason for being sanctioned. The appearance of "trust" and "consent" also tracks, given the nature of the attorney-client relationship. The fact that "account" and "funds" sit so high on the list is interesting; it suggests that sanctions are often for financial impropriety.**

------------------------------------------------------------------------

##### **Q2.** Let's move beyond single words to phrases. Make a list of the top 10 three-word phrases, called trigrams, based on the example from the pre_lab (you'll need to modify the example code to do this). What's the top trigram and how often does it appear? What does that phrase mean in legal terms?

```{r}
# Check the task 12 on Pre-lab 11 to see if you can use that code
trigrams <- md_attorney_sanctions |>
  unnest_tokens(trigram, text, token = "ngrams", n = 3) |>
  separate(trigram, c("word1", "word2", "word3"), sep = " ") |>
  filter(!word1 %in% stop_words$word) |>
  filter(!word2 %in% stop_words$word) |>
  filter(!word3 %in% stop_words$word) |>
  mutate(trigram = paste(word1, word2, word3, sep=" ")) |>
  group_by(trigram) |>
  tally(sort=TRUE) |>
  mutate(percent = (n/sum(n))*100) |>
  top_n(10)

```

**A2. "conduct involving dishonesty," "dishonesty fraud deceit" and "involving dishonesty fraud" are the most frequently used phrases, each accounting for roughly three percent of total words. A "fraudulent" or "dishonest" act can describe withholding information, manipulation or lying – sometimes with the intent to obtain money.**

------------------------------------------------------------------------

##### **Q3.** Let's drop back down to more traditional text analysis - take the top trigram from Q2 and write code to see how many times it occurs in the text column in each fiscal year. What do you think the answer produced by your code suggests? What else could you do to try and clarify the most important reasons attorneys get sanctioned?

```{r}

by_year <- md_attorney_sanctions |> filter(str_detect(text, "conduct involving dishonesty")) |> group_by(fiscal_year) |> summarise (count = n()) |> arrange(desc(count))


```

**A3. It seems that 2019-2022 was a busy time for attorneys conducting themselves dishonestly. Jumping to conclusions, one could look at this and say an increasing number of Maryland attorneys are engaging in dishonest behavior. But there are plenty of other explanations. Maybe the entity in charge of sanctioning has become more efficient. Maybe the total number of lawyers in Maryland has significantly increased, meaning that sanctions – although greater in number – have remained consistent. Or, it could simply be that there's a new official in charge of writing the text of the sanctions. And they prefer "conduct involving dishonesty" over a different phrasing. Whatever the case, more analysis would be required to know for sure. For more precise text analysis, it might be helpful to look at the set of rules governing attorney conduct and see what specific legal terms they use for offences.**
