---
title: "lab_03"
author: "Derek Willis"
date: "2023-02-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

#Load the tidyverse
library (tidyverse)

install.packages ('janitor')
library (janitor)

# Load janitor
```

## Load data and clean up the column names

Load the two datasets we worked with in the pre_lab: the Maryland payments to vendors and Maryland grants and loans and, using janitor, change the columns so that they are lower-cased and any spaces are replaced with underscores.

```{r}
# Load in the Maryland state agency payments data, making sure that it correctly sets the datatypes using all of the rows

# Load in the Maryland grants and loan data

grants_loans <- read_csv ("data/State_of_Maryland_Grant_and_Loan_Data__FY2009_to_FY2022.csv") |>
  clean_names()

payments <- read_csv ("data/State_of_Maryland_Payments_Data__FY2008_to_FY2024.csv") |>
  clean_names ()
```

## Answer questions

Q1: You want to know how much money state agencies have paid to the Prince George's County School system, but to do that you'll need to find all of the variations of how the school system might be listed. Write code to find rows where the vendor name contains the county's school system, starting with a broad search and then narrowing it down. When you have found as many as possible, save the result to a new dataframe called `pg_schools`. How many records did you find?

A1: **66 observations.**

```{r}
pg_schools <- payments |> filter (
  str_detect (vendor_name, "PG COUNTY PUBLIC SCHOOL") |
  str_detect(vendor_name, "PRINCE GEORGES COUNTY SCHOOL") |
  str_detect(vendor_name, "PRINCE GEORGES PUBLIC SCHOOLS") |
  str_detect(vendor_name, "PRINCE GEORGES COUNTY PUBLIC SCHOOL") | 
  str_detect (vendor_name, "PRINCE GEORGES COUNTY PUBLIC SCHOOLS"))

#PG COUNTY PUBLIC SCHOOL, PRINCE GEORGES COUNTY SCHOOL, PRINCE GEORGES PUBLIC SCHOOLS, PRINCE GEORGES COUNTY PUBLIC SCHOOL, PRINCE GEORGES COUNTY PUBLIC SCHOOLS
```

Q2: Using that `pg_schools` dataframe, write code to calculate the total amount of payments to the Prince George's County School system for each fiscal year. Which fiscal year had the largest total amount? What do you notice about the years listed, and what questions do you have about that data?

A2: **Fiscal year 2021. There has been consistent increase in the years since 2018. I'm particularly interested in a sizable increase between 2018 and 2019. What happened between those years to precipitate that budget increase?**

```{r}
pg_schools |> 
  group_by (fiscal_year) |>
  summarise (total_amount = sum(amount)) |>
  arrange (desc(total_amount))
```

Q3. You're interested in examining the vendors paid for capital projects at the University of Maryland. Search the payments data to find those payments, then write code to produce a list of each vendor and the total amount it received. Look carefully at that list - what cleaning will it require to ensure that you have only one row per vendor and zip code? Provide at least three examples of vendor names that are similar and should be combined into one. Describe how you might clean that data so that any nearly identical vendors would look identical - what would you change and how?

**A3.** **The first cleaning steps I would take are removing spaces, commas, and periods from titles like INC and LLC. Examples are a) BRUKER BIOSPIN CORP and BRUKER BIOSPIN CORPORATION b) A&I INC and A&I INC. c) CANNON DESIGN and CANNON DESIGN INC. Besides punctuation changes, I would remove titles like CORP and INC. I would also change all ANDs to &.**

```{r}
capital_umd <- payments |> filter (
  str_detect (agency_name, "CAPITAL PROJECTS- UNIVERSITY OF MARYLAND")) 

capital_umd <- capital_umd |>
  mutate(
    clean_vendor_name = gsub(" ", "", vendor_name), clean_vendor_name = gsub (",", "", clean_vendor_name), clean_vendor_name = gsub ("INC", "", clean_vendor_name), clean_venor_name = gsub("CORPORATION", "", clean_vendor_name), clean_vendor_name = gsub("CORP", "", clean_vendor_name), clean_vendor_name = gsub ("CORP.", "", clean_vendor_name), clean_vendor_name = gsub("AND", "&", clean_vendor_name), clean_vendor_name = gsub("LLC", "", clean_vendor_name)
  )

capital_umd |> group_by (clean_vendor_name) |> 
  arrange (clean_vendor_name) |> 
  summarise (total_amount = sum(amount))
  

```

```{r}

```

Q4: You want to understand how much state grant money went to stem cell research projects each fiscal year. Search the data to find grants that fit that description. Once you find which columns contain the information you want, your code should standardize the data needed to find just the grants for stem cell research in a single filter. Then create a total amount of money for each fiscal year. Write a sentence summarizing how stem cell research funding has changed over time.

**A4: Funding for Stem Cell research seemed to hit its peak between 2009 and 2013, followed by a steep drop off in 2015. Grants picked up significantly in 2019 but have since decreased again.**

```{r}
stem_cell_research <- cleaned_md_grants_loans |> 
  filter(str_detect(description, "STEM CELL"))


```

```{r}
stem_cell_research |> 
  group_by(fiscal_year) |>
  summarise (total_amount = sum(amount)) |>
  arrange(desc(total_amount))
```
